---
title: "P8105_hw3_jl6648"
output: github_document
author: Jixin Li
date: "2023-10-09"
---

# probelm 2 
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(p8105.datasets)                                                                  # load the data
data("brfss_smart2010") 
```

```{r}
brfss_df =
  brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  separate(locationdesc, into = c("state", "county"), sep = " - ") |>
  filter(response == "Excellent" | response == "Very good" | response == "Good" | 
         response == "Fair" | response == "Poor") |>
  mutate(response = as.factor(response),
         response = ordered(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) 
```

## question 1
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
prob1_q1_02 = filter(brfss_df, year == 2002) |>
  group_by(state) |>
  summarize(n_obs = n_distinct(county)) |>
  filter(n_obs >= 7) 

state_2002 = pull(prob1_q1_02, state)
```
In 2002, following states `r state_2002` were observed at 7 or more location

```{r}
prob1_q1_10 = filter(brfss_df, year == 2010) |>
  group_by(state) |>
  summarize(n_obs = n_distinct(county)) |>
  filter(n_obs >= 7) 

state_2010 = pull(prob1_q1_10, state)
```
In 2010, following states `r state_2010` were observed at 7 or more location

## question 2
```{r, message=FALSE}
brfss_excellent =
brfss_df |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarise(avg_data = mean(data_value, na.rm = T))

brfss_excellent |>
ggplot(aes(x = year, y = avg_data, color = state)) +
  geom_path() +
  labs(
    title = "average value over time within a state",
    x = "Year",
    y = "average value",
    caption = "Data from the p8105.datasets package"
  ) +
  theme_minimal()
```

## question 3
```{r}
brfss_NY =
  brfss_df |> 
  filter(state == "NY") |>
  filter(year == 2006 | year == 2010) |>
  select(year, state, county, response, data_value) |>
  group_by(year, state)

brfss_NY |>
ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +  
  facet_wrap(~ year, ncol = 2) +
  labs(
    title = "average value among locations in NY states in 2006 and 2010",
    x = "Year",
    y = "average value",
    caption = "Data from the p8105.datasets package"
  ) 
```

# problem 3
```{r}
covar_df =
  read_csv("nhanes_covar.csv", skip=4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
   mutate(                                  
    sex = replace(sex, sex == 2, "Female"),
    sex = replace(sex, sex == 1, "Male"),
    education = replace(education, education == 1, "Less than high school"),
    education = replace(education, education == 2, "High school equivalent"),
    education = replace(education, education == 3, "More than high school")
   ) |>
  mutate(sex = as.factor(sex),
         education = as.factor(education)
  ) |>
  na.omit()
```

```{r}
accel_df =
  read_csv("nhanes_accel.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    min1:min1440,
    names_to = "number",
    values_to = "counts",
    names_prefix = "min"
  )
```

```{r}
combine_df =
  left_join(covar_df, accel_df, by = "seqn")
```

### Produce a reader-friendly table for the number of men and women in each education category
```{r}
sex_education_table =
  select(covar_df, sex | education) 
  
table(sex_education_table)
```

### create a visualization of the age distributions for men and women in each education category
```{r}
covar_df |>
ggplot(aes(x = age, fill = sex )) +
  geom_density(alpha = .5) +
  facet_grid(. ~ education)
  labs(
    title = "age distributions for men and women in each education category",
    x = "age",
    y = "density",
  ) +
  theme_minimal()
```

### Plot these total activities (y-axis) against age (x-axis)
```{r}
accel_clean = 
  combine_df |>
  group_by(seqn, sex, age, education) |>
  summarize(
    total_activity = sum(counts)
  )

accel_clean |>
ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education)
  labs(
    title = "total activities against age",
    x = "age",
    y = "total activities",
  ) 
```

### Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.
```{r}
combine_df |>
ggplot(aes(x = number, y = counts, group = seqn, color = sex)) +
  geom_line(alpha = .2) +
  facet_grid(. ~ education) +
  geom_smooth(aes(group = sex), se = FALSE)
  labs(
    title = "total activities over 24-hour",
    x = "time(mins)",
    y = "total activities",
  ) 
```


